A robust and memory-efficient transition state search method for complex energy landscapes

### 2022-08-22 manuscript.tex
Added parameters for toy2d potential

\section{Results and Discussion}
\subsection{Comparison with other bracketing methods}
\begin{figure}[htb]
  \includegraphics[width=\linewidth]{bracketing.pdf}
  \caption{\label{fig:bracketing}
    Trajectories of the bracketing methods on a hooked potential with a single saddle point.
    The equation for this potential is provided in \cref{sec:bracketingpotential}.
    The minimum energy pathway is shown by the dashed line.
  }
\end{figure}
The BITSS potential in \cref{eq:bitss} and iterative steps above offer key advantages over existing bracketing methods that also use two states to locate the transition state.
For instance, in the ridge method \cite{Ionova1993}, the two images are initially chosen to bracket the largest energy point on an interpolated path between the two endpoints.
However, this is not guaranteed to be on the ridge containing the transition state, and specialist methods are required to avoid high-energy local maxima, or when the initial path contains multiple candidate maxima.
In another example, the double ended surface walking method \cite{Zhang2013} requires Gaussian bias potentials to be added at each iteration to force two dimers to climb uphill in the landscape.
For high numbers of degrees of freedom and many iterations, this becomes very computationally expensive.

@@

\section{Energy and gradient expressions for the example systems}
\subsection{Hook potential equation}\label{sec:bracketingpotential}
The 2D potential used for the comparison of bracketing methods in \cref{fig:bracketing} uses a sum of Gaussian potentials, \cref{eq:bracketingparams}, the parameters for which are provided in \cref{tab:bracketingparams}.
\begin{equation}
  E(x,y) = \sum_i a_i \exp\!\left( \frac{(x-b_{x,i})^2}{c_{x,i}} + \frac{(y-b_{y,i})^2}{c_{y,i}} \right)
  \label{eq:bracketingparams}
\end{equation}
\begin{table}[ht]
  \begin{ruledtabular}
  \begin{tabular}{ccccc}
    $a$  & $b_x$ & $b_x$ & $c_x$ & $c_y$ \\
    \hline
    -1   &   0   &   0   &  10   &  10   \\
     1   &   0   &   0   &   1   &   1   \\
     5   &   2   &   0   &   1   &  0.1  \\
    -1   &   1   &   1   &  0.1  &  0.1  \\
    -1   &   1   &  -1   &  0.1  &  0.1  \\
    0.01 &   0   &   0   &   1   &   1   \\
    0.5  &  -2   &   0   &   1   &   1   \\
  \end{tabular}
  \end{ruledtabular}
  \caption{\label{tab:bracketingparams}
    Parameters for the Gaussians to produce the potential used in \cref{fig:bracketing}.
  }
\end{table}

>>>>

\section{Results and Discussion}
\subsection{Comparison with other bracketing methods}
\begin{figure}[htb]
  \includegraphics[width=\linewidth]{bracketing.pdf}
  \caption{\label{fig:bracketing}
    Trajectories of the bracketing methods on a hooked potential with a single saddle point.
    The equation for this potential is provided in \cref{sec:2dpotentials}.
    The minimum energy pathway is shown by the dashed line.
  }
\end{figure}
The BITSS potential in \cref{eq:bitss} and iterative steps above offer key advantages over existing bracketing methods that also use two states to locate the transition state.
For instance, in the ridge method \cite{Ionova1993}, the two images are initially chosen to bracket the largest energy point on an interpolated path between the two endpoints.
However, this is not guaranteed to be on the ridge containing the transition state, and specialist methods are required to avoid high-energy local maxima, or when the initial path contains multiple candidate maxima.
In another example, the double ended surface walking method \cite{Zhang2013} requires Gaussian bias potentials to be added at each iteration to force two dimers to climb uphill in the landscape.
For high numbers of degrees of freedom and many iterations, this becomes very computationally expensive.

@@

\section{Energy and gradient expressions for the example systems}
\subsection{2D potentials}\label{sec:2dpotentials}
The 2D potentials in \cref{fig:toy2d,fig:bracketing} use a sum of Gaussian potentials, \cref{eq:gaussianpeaks}, the parameters for which are provided in \cref{tab:toy2dparams,tab:bracketingparams}.
\begin{equation}
  E(x,y) = \sum_i a_i \exp\!\left( - \frac{(x-b_{x,i})^2}{c_{x,i}} - \frac{(y-b_{y,i})^2}{c_{y,i}} \right)
  \label{eq:gaussianpeaks}
\end{equation}

\begin{table}[ht]
  \begin{ruledtabular}
  \begin{tabular}{ccccc}
    $a$ & $b_x$ & $b_y$ & $c_x$ & $c_y$ \\
    \hline
    -3  & -1.4  &   0   &   1   &   1   \\
    -2  &  1.4  &   0   &   1   &   1   \\
    -1  &  0.07 &   1   &   1   &   1   \\
  \end{tabular}
  \end{ruledtabular}
  \caption{\label{tab:toy2dparams}
    Parameters for the Gaussians to produce the potential used in \cref{fig:toy2d}.
  }
\end{table}

\begin{table}[ht]
  \begin{ruledtabular}
  \begin{tabular}{ccccc}
    $a$  & $b_x$ & $b_y$ & $c_x$ & $c_y$ \\
    \hline
    -1   &   0   &   0   &  10   &  10   \\
     1   &   0   &   0   &   1   &   1   \\
     5   &   2   &   0   &   1   &  0.1  \\
    -1   &   1   &   1   &  0.1  &  0.1  \\
    -1   &   1   &  -1   &  0.1  &  0.1  \\
    0.01 &   0   &   0   &   1   &   1   \\
    0.5  &  -2   &   0   &   1   &   1   \\
  \end{tabular}
  \end{ruledtabular}
  \caption{\label{tab:bracketingparams}
    Parameters for the Gaussians to produce the potential used in \cref{fig:bracketing}.
  }
\end{table}
